---
title: "test_example"
author: "Zhe GUAN"
date: "2024-07-07"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## R Markdown

```{r}
options(repos = "https://cran.rstudio.com/")
install.packages("data.table")
install.packages("ggmosaic")
library(ggmosaic)
library(data.table)
install.packages("tidyverse")
install.packages("dplyr")
library(tidyverse)
library(dplyr)
library(lubridate)
```




```{r}
filepath <- "/Users/zheguan/CWR_fig/QVI_case/"
transaction_data <- fread(paste(filepath,"QVI_transaction_data.csv",sep = ""))
customer_data <- fread(paste0(filepath,"QVI_purchase_behaviour.csv"))
str(customer_data)
```

```{r}
transaction_data$DATE <- as.Date(transaction_data$DATE,origin = "1899/12/30")
transaction_data[,.N,PROD_NAME]
```


```{r}
transaction_data$DATE <- as.Date(transaction_data$DATE, origin = "1899-12-30")
transaction_data[,.N,by="PROD_NAME"]
str(transaction_data)
```

```{r}
test <- transaction_data[,PROD_NAME] %>% unique() %>% strsplit(.," ") %>% unlist() %>% data.table()  #使用管道挑选出关键词
setnames(test,"words")

test<-test[grepl("[[:digit:]]",words) == FALSE,]  #使用data.table 挑选出不含数字的行。
test<-test[grepl("[[:alpha:]]",words),]
test[,.N,words][order(N,decreasing = TRUE)]
transaction_data<-transaction_data[,SALSA := grepl("salsa",tolower(PROD_NAME))]
transaction_data<-transaction_data[SALSA==FALSE,][, SALSA := NULL]
summary(transaction_data)
tibble(transaction_data)
transaction_data[PROD_QTY == 200,]
transaction_data[LYLTY_CARD_NBR == 226000,]
transaction_data <- transaction_data[LYLTY_CARD_NBR != 226000,]
summary(transaction_data)
transaction_data[,.N,DATE]
allDate <- data.table(seq(as.Date("2018/07/01"),as.Date("2019/06/30"),by= "day"))
setnames(allDate,"DATE")
test3 <- merge(transaction_data[,.N,DATE],allDate,by = "DATE", all.x = TRUE )
#test3 <- allDate %>% left_join(transaction_data[,.N,DATE])
tibble(test3)
theme_update(plot.title = element_text(hjust = 0.5))

#ggplot(test3,aes(x=DATE,y = N)) + 
#ggplot(test3,aes(x = DATE,y = N)) + geom_line() +
ggplot(test3, aes(x = DATE, y = N)) +
geom_line() +
labs(x = "Day", y = "Number of transactions", title = "Transactions over time") +
scale_x_date(breaks = "1 month") +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```
```{r}
ggplot(test3[month(DATE)==12,],aes(x=DATE,y=N))+geom_line()+labs(x="DATE",y="number of transaction",title="Transaction over time in December.")+scale_x_date(breaks = "1 day")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

summarize the package size from transaction_data:
```{r}
package_size <- transaction_data[,package_size := parse_number(PROD_NAME)]
package_size <- transaction_data[,.N,package_size]
package_size[,][order(-package_size)]
```

the most popular product's package size is 175g while the largest package size on market is 380g.

```{r}
ggplot(transaction_data,aes(x=parse_number(PROD_NAME))) + geom_histogram(fill="blue",color = "blue") + labs(x= "package size", y = "the counts of products on sale") 
?geom_histogram
```




select key words from transaction_data

```{r}
productWords <- data.table(unlist(strsplit(unique(transaction_data[,
PROD_NAME]), " ")))
setnames(productWords,"words" )
tibble(productWords)
print(productWords)
productWords <- productWords[grepl("\\d", words) == FALSE, ]
productWords <- productWords[grepl("[:alpha:]", words), ]
productWords[,.N,words][order(N,decreasing = TRUE)]
```

```{r}
colnames(transaction_data)

```

A example of mosaic plot can be used to easily identify the propotions of sales in different groups.

```{r}
total_data <- merge(transaction_data, customer_data, all.x= TRUE)
colnames(total_data)
test <- total_data[,.(total = sum(TOT_SALES)),.(LIFESTAGE, PREMIUM_CUSTOMER)]
tibble(test)
p <- ggplot(data = test) + 
geom_mosaic(aes(weight = total, x = product(PREMIUM_CUSTOMER, LIFESTAGE),fill = PREMIUM_CUSTOMER)) +
labs(x = "Lifestage", y = "Premium customer flag", title = "Proportion of sales") +
theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

p +geom_text(data = ggplot_build(p)$data[[1]],aes(x= (xmin+xmax)/2, y = ymin/2+ymax/2, label = as.character(paste(round(.wt/sum(.wt),3)*100,"%"))))

```

```{r}
average_data <- total_data[,.(average=sum(PROD_QTY)/uniqueN(LYLTY_CARD_NBR)),.(PREMIUM_CUSTOMER,LIFESTAGE)][order(-average)]
tibble(average_data)

p2<- ggplot(data = average_data) +
  geom_mosaic(aes(weight = average, x = product(PREMIUM_CUSTOMER, LIFESTAGE),fill = PREMIUM_CUSTOMER))+labs(x="lifestage",y="premium customer",title= "propotions of product numbers") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

#print(ggplot_build(p2)$data[[1]])

p2 + geom_text(data = ggplot_build(p2)$data[[1]],aes(x = xmin/2 + xmax/2, y = ymin/2 + ymax/2, label = as.character(paste0(round(.wt/sum(.wt),3)*100,"%")) ))
  

```
we can find that the older and young family tend to buy more chips and there is no big difference in the propotion of premium ranks.

Another bar plot is also availale:

```{r}
p3 <- ggplot( data = average_data ) + geom_bar(position = position_dodge(), aes(weight = average, x= LIFESTAGE, fill = PREMIUM_CUSTOMER)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) + labs(x = "lifestage", y = "average number of product purchased")
print(ggplot_build(p3)$data[[1]])
p3 + geom_text(data = ggplot_build(p3)$data[[1]], aes(x = x, y = y + 0.1*y, label= as.character(paste(round(weight/sum(weight),3)*100,"%") )),size = 2)
```

Similary, we can calculate average price per unit by LIFESTAGE and PREMIUM_CUSTOMER
```{r}
#prepare the data with the data.table

avg_price <- total_data[,.(avg_pri = sum(TOT_SALES)/uniqueN(LYLTY_CARD_NBR)),.(PREMIUM_CUSTOMER,LIFESTAGE) ]
tibble(avg_price)
```
```{r}
p4 <- ggplot(data = avg_price) + geom_mosaic(aes(weight = avg_pri, x = product(LIFESTAGE,PREMIUM_CUSTOMER),fill = LIFESTAGE)) +theme(axis.title.x = element_text(hjust = 0.5 ,vjust = 0.01))
p4 + geom_text(data = ggplot_build(p4)$data[[1]], aes(x=xmin/2+xmax/2, y = ymin/2+ymax/2, label = as.character(paste0(round(.wt/sum(.wt),3)*100,"%"))),size = 3)

```

```{r}
avg_price1 <- total_data[,.(avg_pri1 = sum(TOT_SALES)/sum(PROD_QTY)),.(PREMIUM_CUSTOMER,LIFESTAGE) ]
tibble(avg_price1)
```

```{r}
p5 <- ggplot( data = avg_price1) + geom_bar(position = position_dodge() ,aes(x = LIFESTAGE,weight = avg_pri1, fill=PREMIUM_CUSTOMER)) +theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) + labs(x = "lifestage", y = "average price per product", title = "Average price in different customer groups") 

p5 + geom_text(data = ggplot_build(p5)$data[[1]], aes(x=x,y=1.05*y,label = round(weight,2)),size = 2)
```

t-test for different products, we mainly focus on the average price per product

```{r}
total_data[,price:=TOT_SALES/PROD_QTY]
t.test(total_data[LIFESTAGE %in% c("YOUNG SINGLES/COUPLES","MIDAGE SINGLES/COUPLES") & PREMIUM_CUSTOMER == "Mainstream",price],total_data[LIFESTAGE %in% c("YOUNG SINGLES/COUPLES","MIDAGE SINGLES/COUPLES") & PREMIUM_CUSTOMER != "Mainstream",price],alternative = "greater")
```
we will focus on an important group which we can develop, young singles and couples to look at their differences compared to other groups.

```{r}
young_segment <- total_data[LIFESTAGE == "YOUNG SINGLES/COUPLES" & PREMIUM_CUSTOMER == "Mainstream",]
other_segment <- total_data[LIFESTAGE != "YOUNG SINGLES/COUPLES" | PREMIUM_CUSTOMER != "Mainstream",]

#define a metric which means that a specific brand propotion of chips in a group.

total_young_pro <- young_segment[,sum(PROD_QTY)]
tibble(total_young_pro)
seg_test <- young_segment[,.(targetpropo = sum(PROD_QTY)/total_young_pro),PROD_NAME][order(-targetpropo)]
tibble(seg_test)

total_other_pro <- other_segment[,sum(PROD_QTY)]
seg_test2 <- other_segment[,.(targetpropo1 = sum(PROD_QTY)/total_other_pro),PROD_NAME][order(-targetpropo1)]
tibble(seg_test2)

com_young_other_brand <- merge(seg_test,seg_test2)[,ratio := targetpropo/targetpropo1][order(-ratio)]

tibble(com_young_other_brand)
```



install.packages("tidyverse")

