---
title: "simulated customer analytics of chips"
author: "Zhe GUAN"
date: "2024-06-19"
output:
  html_document: default
  pdf_document: default
---



## Preparation for datasets

Two datasets are provided from [forage link1](https://cdn.theforage.com/vinternships/companyassets/32A6DqtsbF7LbKdcq/QVI_transaction_data.xlsx) and [forage link2](https://cdn.theforage.com/vinternships/companyassets/32A6DqtsbF7LbKdcq/QVI_purchase_behaviour.csv).

**LINK1** is transfromed to csv format by EXCEL first.

## Settings

```{r environment}

options(repos = "https://cran.rstudio.com/")
install.packages("tidyverse")
install.packages("dplyr")
library(tidyverse)
library(dplyr)
library(lubridate)

transaction_data <- read_csv("/Users/zheguan/CWR_fig/QVI_case/QVI_transaction_data.csv")
purchanse_data <- read_csv("/Users/zheguan/CWR_fig/QVI_case/QVI_purchase_behaviour.csv")

```

### Review the data

LYLTY_CARD_NBR is Primary Key for two tables.

## Clean the transaction data

change the decimal to date
```{r}
transaction_data$DATE <- as_date(transaction_data$DATE,origin = "1899/12/30")
```
Check missing values for each colomn:

```{r}
transaction_data %>% 
  summarise(across(everything(),  ~ sum(is.na(.))))
```
No missing value for table transaction_data.



### check if there is inconsistency for product price for single products with the same product name.

add a new coloum which indicates the unit_price(single_price) for different products.

```{r}
test <- transaction_data %>% 
  mutate(single_price = round(TOT_SALES/PROD_QTY,3)) %>% 
  group_by(PROD_NAME,single_price) %>% 
  summarise(n()) %>% 
  arrange() 
tibble(test)
```
We found that there are some discripency that some records have different unit/single price but belong to same categories/product. we checked them separately.

We first define a function which will return the mode value for different products transactions.

```{r}
get_mode <- function(x) {
  uniqx <- unique(x)
  uniqx[which.max(tabulate(match(x, uniqx)))]
}
```

then check the mode prices for different products:

```{r}
transaction_data_test <- transaction_data %>% 
    mutate(single_price = round(TOT_SALES/PROD_QTY,3))
price_modes <- transaction_data_test %>%
  group_by(PROD_NAME) %>%
  summarise(price_mode = get_mode(single_price))
View(price_modes)
```
Here are some products which are sold without normal price, so we need to report them for data specialists to check the specific cases.

They are "Dorito Corn Chp Supreme 380g", "Doritos Corn Chips Cheese Supreme 170g", "Grain Waves Sweet Chilli 210g", "Grain Waves Sour Cream&Chives 210G","Infuzions BBQ Rib Prawn Crackers 110g","Kettle Chilli 175g","Kettle Original 175g","Kettle Sensations Camembert & Fig 150g", "Kettle Sweet Chilli And Sour Cream 175g", "Kettle Tortilla ChpsFeta&Garlic 150g", "Old El Paso Salsa Dip Chnky Tom Ht300g", "Old El Paso Salsa Dip Tomato Mild 300g", "Pringles Original Crisps 134g", "Pringles Sthrn FriedChicken 134g", "RRD SR Slow Rst Pork Belly 150g", "Red Rock Deli Thai Chilli&Lime 150g", "Smiths Crinkle Cut Chips Chicken 170g, "Thins Chips Salt & Vinegar 175g", "Tyrrells Crisps Ched & Chives 165g".  We need to report these records for further checks to ensure they are normal cases.


Based on business questions, we only focus on the general cases where products were sold with original price. So we filter transaction_data_test based on the calculated mode:

```{r}
transaction_data_test_filtered <- transaction_data_test %>%
  left_join(price_modes, by = "PROD_NAME") %>%
  filter(single_price >= price_mode)
```

Double Check:

```{r}
test <- transaction_data_test_filtered %>%  
  group_by(PROD_NAME,single_price) %>% 
  summarise(n()) %>% 
  arrange() 
tibble(test)
```




## Clean the purchase data

check the missing values:
```{r}
purchanse_data %>% summarise(across(everything(), ~ sum(is.na(.))))
```

check duplicate values for card number:

```{r}
purchanse_data %>% 
  group_by(LYLTY_CARD_NBR) %>% 
  summarise(n()) 
```

or 

```{r}
purchanse_data %>% 
  group_by(LYLTY_CARD_NBR) %>% 
  summarise(number=n()) %>% 
  filter(number != 1)
```

we found that the card number type in the purchanse_data table is different from transaction table, we need to transfer one of them.

```{r}
purchase_data_cleaned <- purchanse_data %>% 
  mutate(LYLTY_CARD_NBR = as.integer(LYLTY_CARD_NBR))
View(purchase_data_cleaned)
```

## link the purchase table and transction table


```{r}
total_table <- transaction_data_test_filtered %>% 
  left_join(purchase_data_cleaned, by = "LYLTY_CARD_NBR")
View(total_table)
```


## define the metrics 

By investigate the database, we found that we can compare different patterns from different levels: total_sales/total_quantity vs. lifestage/premium_customer/date.

first lifestage versus total sales and quantities.

```{r}
summary_table_lifestage <- total_table %>% 
  group_by(LIFESTAGE) %>% 
  summarise(total_sales = sum(TOT_SALES),total_quantity = sum(PROD_QTY),average_sales=sum(TOT_SALES)/n())
tibble(summary_table_lifestage)
```

in the first step, we can find the **New family** generally tends to buy less chips compared to other kinds of families and **Older families** and **Retirees** tend to buy more chips compared to others. We can recommend stakeholders to focus on these two kinds of groups to organise possible promotional campaigns. Need to mention that Midage Singles and couples are also possible customers with higher average spendings on chips.

More details can be viewed in plot:

```{r}
ggplot(data = summary_table_lifestage, aes(x = LIFESTAGE, y = total_sales,  fill = LIFESTAGE ) ) + geom_col() +labs(title = "The sales vs. lifestage of customers", x = "life stage") + theme(axis.text.x = element_text( size = 4.5, angle = 25 ))
```

similarily, we can make more plots for different ranks of customers, and in this case, we will focus on the average sales:

```{r}
summary_table_premium <- total_table %>% 
  group_by(PREMIUM_CUSTOMER) %>% 
  summarise(total_sales = sum(TOT_SALES),total_quantity = sum(PROD_QTY), average_sales = sum(TOT_SALES)/n())
tibble(summary_table_premium)
```
More details can be viewed in plot:

```{r}
ggplot(data = summary_table_premium, aes(x = PREMIUM_CUSTOMER, y = total_sales,  fill = PREMIUM_CUSTOMER ) ) + geom_col() +labs(title = "The sales vs. lifestage of customers", x = "Ranks") + theme(axis.text.x = element_text( size = 10, angle = 0 ))
```

just look at the total sales, **Mainstream** contributes most and next one is Budget. Budget is the next one, and Premium contribute least. However, if we look at the average sales, there is no big difference between Budget and Premium. So we can suggest stakeholders focus on the Mainstream market. 

Next business question is : Which chip kind is the most popular?

First, we grouped the dataset based on product_name/numbers:

```{r}
total_table_chips <- total_table %>% 
  group_by(PROD_NBR,PREMIUM_CUSTOMER,PROD_NAME) %>%
  summarise(total_quantity=sum(PROD_QTY),total_sales=sum(TOT_SALES),average_sales=sum(TOT_SALES)/sum(PROD_QTY))
tibble(total_table_chips)
```

Then check them in three different premium of customers:

for MainStream:

```{r}
total_table_chips_mainstream <- total_table_chips %>% 
  filter(PREMIUM_CUSTOMER == "Mainstream") %>% 
  arrange(desc(total_sales)) %>% head(10)
tibble(total_table_chips_mainstream)
ggplot(total_table_chips_mainstream) + geom_col(mapping = aes(x=total_sales,y = PROD_NAME ,fill = PROD_NAME)) + labs(title = "Top 10 Products in Mainstream Market", subtitle = "Based on total sales from 2018") +theme(title = element_text(size = 9)) + guides(fill = FALSE) 

```

For Budget: 

```{r}
total_table_chips_budget <- total_table_chips %>% 
  filter(PREMIUM_CUSTOMER == "Budget") %>% 
  arrange(desc(total_sales)) %>% head(10)
tibble(total_table_chips_budget)
ggplot(total_table_chips_budget) + geom_col(mapping = aes(x=total_sales,y = PROD_NAME ,fill = PROD_NAME)) + labs(title = "Top 10 Products in Budget Market", subtitle = "Based on total sales from 2018") +theme(title = element_text(size = 9)) + guides(fill = FALSE) 
```

For Premium:

```{r}
total_table_chips_premium <- total_table_chips %>% 
  filter(PREMIUM_CUSTOMER == "Premium") %>% 
  arrange(desc(total_sales)) %>% 
  head(10)
tibble(total_table_chips_premium)
ggplot(total_table_chips_premium) + geom_col(mapping = aes(x=total_sales,y = PROD_NAME ,fill = PROD_NAME)) + labs(title = "Top 10 Products in Premium Market", subtitle = "Based on total sales from 2018") +theme(title = element_text(size = 9)) + guides(fill = FALSE) 

```

We can found the most popular product in three groups is **Dorto Corn Chp Supreme 380g**. Perhaps its some characters make it stand out. And other products can be improved based on it.

We also investigate the Old Singles/Couples have their own preferences because they contribute most in total sales.
```{r}
total_table_chips_life <- total_table %>% 
  group_by(PROD_NBR,LIFESTAGE,PROD_NAME) %>%
  summarise(total_quantity=sum(PROD_QTY),total_sales=sum(TOT_SALES),average_sales=sum(TOT_SALES)/sum(PROD_QTY))
tibble(total_table_chips_life)
```

```{r}
total_table_chips_premium <- total_table_chips_life %>% 
  filter(LIFESTAGE == "OLDER SINGLES/COUPLES") %>% 
  arrange(desc(total_sales)) %>% 
  head(10)
tibble(total_table_chips_premium)
ggplot(total_table_chips_premium) + geom_col(mapping = aes(x=total_sales,y = PROD_NAME ,fill = PROD_NAME)) + labs(title = "Top 10 Products for OLDER SINGLES/COUPLES", subtitle = "Based on total sales from 2018") +theme(title = element_text(size = 8)) + guides(fill = FALSE) 
```

the most popular product is still **Dorito Corn Chp Supreme 380g**.

We can do further check to check the correlations between the most popular product/groups and sales and need more time for that.

How the sales change with the date? 

```{r}
total_table_date <- total_table %>% 
  group_by(DATE) %>%  
  summarise(total_quantity=sum(PROD_QTY),total_sales=sum(TOT_SALES),average_sales=sum(TOT_SALES)/sum(PROD_QTY))
ggplot(total_table_date) + geom_col(mapping = aes(x=DATE,y=total_sales))
```

We can easily found that nealy November and December, the chips sales increased a lot, but during the **mid-May** to **mid-June**, the sales would experieced regular drop.

We suggest stakeholders to prepare/organize possible sales promotion in Nov. to Dec.




## Summary and suggestions

* **Older families** and **Retirees** contribute most in total chips sales. New families purchase least but also could be a possible market to dip.

* For different ranks of customers, **Mainstream** is the most important part, and we recommend to focus on Mainstream market. Although total sales of customers with "Budget" is larger than that of customers with "Premium", the average sales is pretty close.

* Although there are some fluctuations for top 10 popular chips, the most popular product in three groups is **Dorto Corn Chp Supreme 380g** without doubt. Perhaps its some special characters make it stand out. And other products can be improved based on it.

* The data also reveals that the sales is increased during **November** to **December** and the lowest vales appears in **May**. We recommend to focus on the time point to prepare promotion activities.

NOTES:
* Although different metrics used(total_sales/total_quantity/average_sales) showed similar patterns, more details of products/dates could be studied based on correlations among different groups.

* The sales characteristics is based on data between 2018 and 2019, more data should be reviewed in case some variations.


